---
title: 'Is the cloud cost-effective?'
author: "Steph Locke (@SteffLocke)"
date: '`r Sys.Date()`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r autodoc, child='AboutMeTech.Rmd', eval=TRUE}
```

## Agenda
- Useful concepts
- Costs of being on-prem
- Costs of being in the cloud
- Building a financial case


# Useful concepts

## Compound interest

## Depreciation

## Opportunity costs

# Costs of being on-prem
## On-prem new
- Device
- Software
- Implementation time
- Ongoing costs
   + Power / datacentre costs
   + Staff
   + Support agreements
   + Backups

## On-prem upgrade
- Device
- Software
- Migration time
- Ongoing costs
   + Power / datacentre costs
   + Staff
   + Support agreements
   + Backups

# Costs of being in the cloud
## Cloud new
- Ongoing cost
- Implementation time

## Cloud migration
- Ongoing cost
- Migration time

# Building a financial case
## Cost of development
- How many days to upgrade?
- How many days to convert?
- How many days to do like-for-like migration?
- What are the typical costs?
- What won't we be able to do whilst we're doing the development?

## Physical costs
- What are the costs for hardware?
- How long will the hardware be in use?
- Are we fitting to current requirements or requirements towards EOL?
- What are the licence costs?
- What are the ongoing costs of maintaining the server?

## Cloud costs
- What are the costs currently?
- What are the growth assumptions?
- Are there change points in cost that will be encountered?

## Demo
[App](https://stef.shinyapps.io/is_the_cloud_cost_effective/)

# Wrapup
## Wrapup
- Thank you
- Get the stuff
- Get in touch! [T: SteffLocke](https://twitter.com/stefflocke)
