---
title: 'Your first bot - no code required!'
author: "Steph Locke (@SteffLocke)"
date: '`r Sys.Date()`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r autodoc, child='AboutMeTech.Rmd', eval=TRUE}
```

## Agenda
- Aim
- Demo
- Theory
- Recommended resources

# Build a bot without writing code {.quote}

# Demo time! {.quote}

# Theory
## QnAMaker.ai
QnAMaker provides a knowledge base system for Q&As.

- Basic processing
- Active Learning

[qnamaker.ai](//qnamaker.ai)

## Bot Builder SDKs
Code in C# and node.js for working with bots

[gh:microsoft/botbuilder](//github.com/Microsoft/BotBuilder)

## Bot Services
Function based platform for making bot APIs

[docs.microsoft.com](//docs.microsoft.com/en-gb/bot-framework/azure-bot-service-overview)

## Bot Framework
Bot management area

[botframework.com](//botframework.com)

## Channels
Ways people can consume your bots

## LUIS
LUIS allows you to manage concepts and decode people's text

[luis.ai](//luis.ai)

# Recommended resources

## Recommended resources
- [Botscamp](//www.botscamp.co)
- GitHub
- [Videos of demo](//youtube.com/playlist?list=PL1ROw3K8pbB2nogCjfQYSMNzTwn_fReWc)
- [Channel 9](//channel9.msdn.com/Search?term=bot#ch9Search&lang-en=en&pubDate=year)

# Wrapup
## Wrapup
- Thank you for attending
- Thank you to the folks who made this happen
- Get in touch [\@SteffLocke](https://twitter.com/stefflocke)
