---
title: 'Your first bot - no code required!'
author: "Steph Locke (@SteffLocke)"
date: '`r Sys.Date()`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r autodoc, child='AboutMeTech.Rmd', eval=TRUE}
```

## Agenda
- Aim
- Demo
- Theory
- Recommended resources

# Build a bot without writing code {.quote}

# Demo time! {.quote}

# Theory
## QnAMaker.ai
QnAMaker provides a knowledge base system for Q&As.

- Basic processing
- Active Learning

## Bot Builder SDKs
Code in C# and node.js for working with bots

## Bot Services
Function based platform for making bot APIs

## Bot Framework
Bot management area

## Channels
Ways people can consume your bots

## LUIS
LUIS allows you to manage concepts and decode people's text

# Recommended resources

## Recommended resources
- Botcamp
- GitHub
- Videos of demo
- Channel 9

# Wrapup
## Wrapup
- Thank you for attending
- Thank you to the folks who made this happen
- Get in touch [\@SteffLocke](https://twitter.com/stefflocke)
