---
title: 'R & SQL Server for real-time predictions'
author: "Steph Locke\n@SteffLocke\nLocke Data"
date: '`r Sys.Date()`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=TRUE, size="large", tidy.opts=list( width.cutoff=60))

do_eval<-FALSE
```

```{r autodoc, child='AboutMeTech.Rmd', eval=TRUE}
```




# R

## Overview
> R is an integrated suite of software facilities for data manipulation, calculation and graphical display

- Open source
- In-memory & single-core (by default)
- Multi-platform
- Extensible environment
- Delivered by the [R Foundation](https://www.r-project.org/foundation/), supported by the [R Consortium](https://www.r-consortium.org/), grown by R developers 
- [r-project.org](http://www.r-project.org/)

## What's it look like?
```{r}
ggplot2::qplot(mpg, disp, data = mtcars, colour = cyl)
```

## Versions
- ["Standard" R](https://r-project.org)
- [Microsoft R Open](https://mran.revolutionanalytics.com/download/)
- [Microsoft ML Server](https://msdn.microsoft.com/en-us/library/mt674876.aspx)
- [TERR](https://docs.tibco.com/products/tibco-enterprise-runtime-for-r)
- [Oracle R](http://www.oracle.com/technetwork/database/database-technologies/r/r-distribution/overview/index.html)

## Key things to manage
- Packages
- RAM
- Code

## Next steps
- [Learn R](https://datacamp.com)
- [Check out some of my other intros to R](http://itsalocke.com/talks)

# SQL Server + R Services

## Overview
- Use `sp_execute_external_script` to call R from SQL
- Store model objects in SQL Server
- Use certain models in a native PREDICT function

## 2016/2017 Editions
- Express (w/ Advanced tools) - R Open or base R
- Standard - R Open or base R
- Enterprise - ML Server

[SQL Server feature comparison](https://docs.microsoft.com/en-us/sql/sql-server/editions-and-components-of-sql-server-2017#Programmability)


## Key things to manage
- RAM & CPU
- Licenses vs Server resources
- Permissions
- Code

## Next steps
- [Tomaz Kastrun blog](https://tomaztsql.wordpress.com/)
- [BOL](https://msdn.microsoft.com/en-us/library/mt604845.aspx)

## Connecting to our demo env
```{r eval=do_eval}
library(DBI)
library(odbc)

driver   = "ODBC Driver 13 for SQL Server"
server   = "sqlsat666.database.windows.net"
database = "sqlsat666"
uid      = "lockedata"
pwd      = "zll+.?=g8JA11111"

dbConn<-dbConnect(odbc(),
          driver=driver, server=server,
          database=database, uid=uid,
          pwd=pwd)
```

```{r echo=FALSE}
if(!exists("dbConn")) dbConn<-dbConnect(ANSI())
```

## Connecting to our demo env (extra)
```{r eval=TRUE}
library(RODBCext)

dbstring <- glue::glue('Driver={driver};Server={server};Database={database};Uid={uid};Pwd={pwd}')
dbconn <- RODBC::odbcDriverConnect(dbstring)
```

```{r echo=FALSE}
if(!exists("dbconn")) dbconn<-dbConnect(ANSI())
```

## A basic execution
```{sql connection=dbConn, eval=do_eval}
EXECUTE sp_execute_external_script
        @language = N'R'
        ,@script = N'OutputDataSet <- InputDataSet'
        ,@input_data_1 = N'SELECT 1 as Col'
 WITH RESULT SETS ((col varchar(50) not null))        
```

# DEMO PREP 

## Setup model storage area

```{sql, connection=dbConn, eval=do_eval, error=TRUE}
CREATE TABLE [companyModels]    (  
  [id] int NOT NULL PRIMARY KEY IDENTITY (1,1)   
, [name] varchar(200) NOT NULL      
, [modelObj] varbinary(max)    
, [ValidFrom] datetime2 (2) GENERATED ALWAYS AS ROW START  
, [ValidTo] datetime2 (2) GENERATED ALWAYS AS ROW END  
, PERIOD FOR SYSTEM_TIME (ValidFrom, ValidTo)  
, CONSTRAINT unique_modelname UNIQUE ([name]))
WITH (SYSTEM_VERSIONING = ON (HISTORY_TABLE = dbo.companyModelsHistory));  
```

## Model UPSERT stored procedure
```{sql, connection=dbConn, eval=do_eval, echo=FALSE, result="hide"}
DROP PROCEDURE IF EXISTS modelUpsert
```

```{sql, connection=dbConn, eval=do_eval}
CREATE PROCEDURE modelUpsert
@modelname  varchar(200) , 
@modelobj varbinary(max) 
AS
WITH MySource as (
    select @modelname as [name], @modelobj as [modelObj]
)
MERGE companymodels AS MyTarget
USING MySource
     ON MySource.[name] = MyTarget.[name]
WHEN MATCHED THEN UPDATE SET 
    modelObj = MySource.[modelObj]
WHEN NOT MATCHED THEN INSERT
    (
        [name], 
        modelObj
    )
    VALUES (
        MySource.[name], 
        MySource.modelObj
    );
```

## Add some data
```{r, echo=FALSE}
dbWriteTable(dbConn, "flights", nycflights13::flights[1:5000,], overwrite=TRUE)
```

```{r, eval=FALSE}
dbWriteTable(dbConn, "flights", nycflights13::flights, overwrite=TRUE)
```


# R DEMO 

## Produce a basic model
```{r}
library(nycflights13)
flightLM<- lm(arr_delay ~ month + day +  hour, data=flights, model = FALSE)

flightLM
```

## Use model in R
```{r}
sample_flights<- flights[1:5,]
predict(flightLM, sample_flights)
```

## Publish model to Microsoft ML Server
[mrsdeploy](https://docs.microsoft.com/en-us/machine-learning-server/operationalize/how-to-deploy-web-service-publish-manage-in-r)

```{r eval=FALSE}
publishService(
  "basicFlightsLM",
  code = NULL,
  # --- `model` is required for web service with serviceType `Realtime` --- #
  model = "model = flightsLM",
  # --- `serviceType` is required for this web service --- #
  serviceType = "Realtime"
)
```

## Store a model
```{r eval=do_eval}
RODBCext::sqlExecute(dbconn, "exec modelUpsert @modelname=? , @modelobj=?",
                     data = data.frame("modelFromR", paste0( serialize(flightLM,NULL)
       ,collapse = "")))
```

# IN-SQL DEMO - ANY MODEL 

## Produce a basic model
```{sql, connection=dbConn, eval=do_eval, echo=FALSE, result="hide"}
DROP PROCEDURE IF EXISTS generate_flightlm
```

```{sql, connection=dbConn, error=TRUE, eval=do_eval}
CREATE PROCEDURE generate_flightlm
AS
BEGIN
EXECUTE sp_execute_external_script
        @language = N'R'
        ,@script = N'
        flightLM<-lm(arr_delay ~ month + day +  hour, data=InputDataSet, model=FALSE)
        OutputDataSet<-data.frame(modelname="modelFromInSQL",
        modelobj=paste0(                              serialize(flightLM,NULL)
       ,collapse = "") )
        '
        ,@input_data_1 = N'SELECT * FROM flights'
       WITH RESULT SETS (([name] varchar(200) NOT NULL   
, [modelObj] varbinary(max) ))
END
```

## Produce a basic model
```{sql, connection=dbConn, error=TRUE, eval=do_eval}
INSERT INTO companyModels(name, modelObj)
EXEC generate_flightlm
```

## Use model in SQL
```{sql, connection=dbConn, eval=do_eval}
DECLARE @mymodel VARBINARY(MAX)=(SELECT modelobj 
                FROM companymodels 
                WHERE [name]='modelFromR'
                );
EXEC sp_execute_external_script
@language = N'R',  
@script = N'
OutputDataSet<-data.frame( predict(unserialize(as.raw(model)), InputDataSet),
InputDataSet[,"arr_delay"]
    )
',
@input_data_1 = N'SELECT TOP 5 * from flights',  
@params = N'@model varbinary(max)',  
@model =  @mymodel 
WITH RESULT SETS ((
    [arr_delay.Pred]  FLOAT (53)    NULL,
    [arr_delay]  FLOAT (53)    NULL))
```

# NATIVE SQL DEMO 
## Produce a basic model
```{sql, connection=dbConn, eval=do_eval, echo=FALSE, result="hide"}
DROP PROCEDURE IF EXISTS generate_flightlm2
```

```{sql, connection=dbConn, error=TRUE, eval=do_eval}
CREATE PROCEDURE generate_flightlm2
AS
BEGIN
EXECUTE sp_execute_external_script
        @language = N'R'
        ,@script = N'
        flightLM<-rxLinMod(arr_delay ~ month + day +  hour, data=InputDataSet)
        OutputDataSet<-data.frame(modelname="modelFromRevo",
        modelobj= rxSerializeModel(flightLM, realtimeScoringOnly = TRUE) )
        '
        ,@input_data_1 = N'SELECT * FROM flights'
END
```

## Produce a basic model
```{sql, connection=dbConn, error=TRUE, eval=do_eval}
INSERT INTO companyModels([name],[modelobj])
EXEC generate_flightlm2
```

## Use model in SQL
```{sql, connection=dbConn, error=TRUE, eval=do_eval}
DECLARE @model varbinary(max) = (
  SELECT modelobj
  FROM companyModels
  WHERE [name] = 'modelFromRevo');
SELECT TOP 10 d.*, p.*
  FROM PREDICT(MODEL = @model, DATA = flights as d)
  WITH(Pred float) as p;
```

## Use model in SQL
```{sql, connection=dbConn, error=TRUE, eval=do_eval}
DECLARE @model varbinary(max) = (
  SELECT modelobj
  FROM companyModels
  WHERE [name] = 'modelFromRevo');

EXEC sp_rxPredict
@model = @model,
@inputData = N'SELECT TOP 10 * FROM flights'
```

# Deploying SQL Server & R Services

## Azure is your friend
- Automatable
- Disposable
- Scalable

## ARM templates
- JSON specification of resources
- Paramaterised for reuse
- Downloadable
- Deploy in many different languages

## PowerShell
- Azure modules
- Use ARM templates
- Used in Azure Automation

## Demo
[Github](https://github.com/stephlocke/SQLServerVM-tidyverse)

## Next steps
- [Authoring ARM templates](https://azure.microsoft.com/en-gb/documentation/articles/resource-group-authoring-templates/)
- [PowerShell basics](https://msdn.microsoft.com/en-us/powershell/scripting/getting-started/fundamental/windows-powershell-basics)
- [Azure overview](https://azure.microsoft.com/en-gb/overview/what-is-azure/)

# Other Microsoft & R things

## Azure ML
- GUI data science
- R & Python
- Notebooks
- Code deployment potential
- Easy webservices

## R in PowerBI
- Interactive reports
- R data source
- R graphics

# Wrapup
## Microsoft resources
- https://docs.microsoft.com/en-us/sql/advanced-analytics/r/sql-server-r-services
- https://docs.microsoft.com/en-us/sql/advanced-analytics/r/how-to-do-realtime-scoring#native-scoring-with-predict
- https://blogs.msdn.microsoft.com/sqlserverstorageengine/2017/09/25/announcing-preview-of-machine-learning-services-with-r-support-in-azure-sql-database/

## Wrapup
- Thank you
- Get the slides via [itsalocke.com](http://itsalocke.com/talks)
- Get in touch! [\@SteffLocke](https://twitter.com/stefflocke)
